The first high-redshift Changing Look Quasars

### 2019-11-07 tex/other_tex/201910/CIV_CLQs_20191021.tex
CIV tex/ tidy up

\citet{Calderone2017} present the Quasar Spectral Fitting (QSFIT)
software package which among other quantities provides luminosity
estimates as well as width, velocity offset and equivalent width of 20
emission lines, includinng \civ, \ciii and \mgii.  We attempt to process and fit
all nine spectra using the lastest version (v1.3.0) of the QSFIT
\href{https://qsfit.inaf.it/cat_1.30/onlinefit.php}{online calculator}
and setting $E(B-V)=0.00$.

@@

 When using the \citet{Shen2011} catalogue, we report line luminosity,
FWHM, rest-frame equivalent width, and their errors for the {\it
whole} \mgii profile (this catalogue also quotes values for just the
broad \mgii compoment, but the differences for our objects is
negligible). The line luminosites in \citet{Calderone2017} are given
in units of $10^{42}$ erg s$^{-1}$, while the line luminosites in
\citet{Shen2011} are in units of $\log(L)$.  We do not report
the \citet{Shen2011} error on the line luminosity.

@@

In the context of CLQs at lower-$z$...  
(Balmer CLQs vs. \civ CLQs...) 
Sociis natoque penatibus et
magnis dis parturient montes, nascetur ridiculus mus. Duis tempus,
lectus nec ultricies mollis, mi orci feugiat nulla, a bibendum velit
orci nec lacus. Duis a odio in nisi egestas dictum. Nullam vel quam
mauris, eget consectetur orci. Morbi ac mi sit amet neque consectetur
tempus ac eget est. Curabitur malesuada arcu sit amet metus dictum at
dapibus arcu accumsan. Fusce sollicitudin luctus rutrum.

@@

\begin{table*}
  \begin{tabular}{ll l l cc ll}
    \hline
    \hline
    Object        & MJD         & $M_i$   &  $L_{\rm bol}$             & \multicolumn{2}{c}{$M_{\rm BH}$}          & $\eta_{\rm Edd}$  & Ref. \\
                      &                &              &                                   & \mgii      & \civ                                    &                             &        \\
    \hline
                      & 53498    & -27.74   & 47.216$\pm$0.004   &  9.55$\pm$0.05   &   9.49$\pm$0.04   &  -0.3783               & Shen2011\\
 J1205+3422  & 58538    &               &                              &                             &                                &                              &  \\
                      & 58693    &               &                              &                            &                                 &                             &  \\
    \hline 
                      & 54553    & -26.75   & 46.166$\pm$0.04    &   9.03$\pm$0.366 & 8.742$\pm$0.15   &  -0.6768               & Shen2011\\
 J1638+2827 & 55832    &  -26.40   & 46.721$\pm$0.073  &   9.338                   &  9.1289                   & -0.717                           & Kol017\\
                      & 58583    &              &                                   &                              &                                 &                   & \\
    \hline 
                      & 56189    &  -25.46  & 46.231$\pm$0.073  &   9.448                 & 8.733                        &  -1.317               & Kol017\\
J2228+2201  & 56960    &               &                                  &                           &                                    &                 &   \\
                      & 58693    &              &                                   &                            &                                   &               &  \\
    \hline
    \hline
  \end{tabular}
  \caption{$\eta_{\rm Edd}$ is the base 10 logarithm of the Eddington ratio.
        Shen11 is \citet{Shen2011}. 
        Kol17 is \citet{Kozlowski2017}.
        Measured line values using the methods and catalogue from
\citet{Shen2011}.  $M_{\rm I}(z=2)$ is the Absolute $i$-band magnitude
$K$-corrected to $z = 2$; Bolometric luminosity computed from the
monochromatic luminosity at 1350\AA\ using the spectral fits and
bolometric corrections (BC = 3.81) in \citet{Richards2006b}; Line
luminosity, FWHM, rest-frame equivalent width, and their errors for
the whole \civ\ profile.  Power-law slope $\alpha_{\lambda}$ for the
continuum fit for \civ; Virial BH masses using calibrations of
\citet{VestergaardPeterson2006}.  Eddington ratio computed using the
virial BH mass.
      }
\end{table*}

@@

\subsection{The Baldwin Effect}
The variable properties of the rest-frame UV quasar emission lines
have been long studied, with the global (or ensemble) Baldwin Effect
(the anti-correlation between the EW of the emission line and the
underlying continuum luminosity of single-epoch observations of a
large number of AGN, first noted in \citet{Baldwin1977}.  More
recently, the intrinsic Baldwin effect, the same anti-correlation but
in an individual, variable AGN.  The variable properites of high-$z$
quasars, however, has been less well stuided. Reasons for this include
more massive systems will tend to have longer timescales e.g. with
$R_{\rm Sch}= 2GM /c^{2}$ and $t_{\rm dyn} =\sqrt{ R^{3}/GM}$ means
$t_{\rm dyn} \sim 6 G^{3} M^{3} / GM c^{2} \sim 6G^{2} M^{2} / c^{2}
$.

\subsection{Eddington ratios and State Changes} 
The broad UV and optical lines in quasars are most sensitive to the extreme ultraviolet (EUV)
part of the spectral energy distribution (SED), with \civ\ (and indeed \heii and \nv) being at
the higher eneggy end of the EUV distribution. 
\citet{NodaDone2018} note that The soft X-ray excess produces most of
the ionizing photons, so its dramatic drop leads to the disappearance
of the broad-line region, driving the ``changing-look'' phenomena.
major difference is that radiation pressure should be much more
important in AGNs, so that the sound speed is much faster than
expected from the gas temperature.
%%
This spectral hardening appears similar to the soft-to-hard state
transition in black hole binaries at $L / L_{\rm Edd} \sim 0.02$
(i.e. $\eta_{\rm Edd} \sim -1.7$, where the inner disc evaporates into
an advection dominated accretion flow, while the overall drop in
luminosity appears consistent with the hydrogen ionization disc
instability.  Crucially \citet{NodaDone2018} make the prediction that
all changing-look AGNs are similarly associated with the state
transition at $L / L_{\rm Edd} \sim$a few per cent.

@@

Etiam mollis viverra nisi eget aliquet. Aliquam erat volutpat. Vivamus
tristique, nisl eu malesuada semper, libero tortor convallis elit, a
scelerisque orci nisi lacinia turpis. In lacinia ultrices
volutpat. Proin ultrices luctus tellus, in placerat eros tincidunt
id. Ut varius iaculis quam in consequat. Nulla nec orci est, sit amet
pellentesque nisl. Mauris non cursus lectus. Praesent placerat leo vel
erat gravida lacinia. Donec vehicula consectetur lectus vitae
luctus. Praesent nisl justo, laoreet elementum facilisis vel,
tristique ac enim. Etiam vel quam ut quam eleifend
tincidunt. Suspendisse sit amet eros vel elit ullamcorper
laoreet. Etiam venenatis sodales turpis, nec lacinia ligula hendrerit
nec. Nam eu vulputate purus. Quisque facilisis congue metus, sed
imperdiet lorem rhoncus sit amet.

Aliquam ac metus nec odio tempus pharetra sed nec diam. Sed eget arcu
nulla. Etiam elementum ultrices ligula, at iaculis libero feugiat
bibendum. Suspendisse potenti. Nam pharetra adipiscing
euismod. Quisque imperdiet dignissim odio, sed volutpat justo
tincidunt eu. Nunc vehicula pharetra suscipit. Integer aliquet pretium
ipsum vel ultrices. Nam rutrum nibh ac quam pulvinar molestie.

>>>>

\citet{Calderone2017} present the Quasar Spectral Fitting (QSFIT)
software package which among other quantities provides luminosity
estimates as well as width, velocity offset and equivalent width of 20
emission lines, including \civ, \ciii and \mgii.  We attempt to process and fit
all nine spectra using the lastest version (v1.3.0) of the QSFIT
\href{https://qsfit.inaf.it/cat_1.30/onlinefit.php}{online calculator}
and setting $E(B-V)=0.00$.

@@

 When using the \citet{Shen2011} catalogue, we report line luminosity,
FWHM, rest-frame equivalent width, and their errors for the {\it
whole} \mgii profile (this catalogue also quotes values for just the
broad \mgii component, but the differences for our objects is
negligible). The line luminosities in \citet{Calderone2017} are given
in units of $10^{42}$ erg s$^{-1}$, while the line luminosities in
\citet{Shen2011} are in units of $\log(L)$.  We do not report
the \citet{Shen2011} error on the line luminosity.

@@

In the context of CLQs at lower-$z$...  
(Balmer CLQs vs. \civ\ CLQs...) 
Sociis natoque penatibus et
magnis dis parturient montes, nascetur ridiculus mus. Duis tempus,
lectus nec ultricies mollis, mi orci feugiat nulla, a bibendum velit
orci nec lacus. Duis a odio in nisi egestas dictum. Nullam vel quam
mauris, eget consectetur orci. Morbi ac mi sit amet neque consectetur
tempus ac eget est. Curabitur malesuada arcu sit amet metus dictum at
dapibus arcu accumsan. Fusce sollicitudin luctus rutrum.

@@

%%
%% %% \eta_{\rm Edd} = log10((10**LBol)/((1.26e38)*(10**MBH)))
%%
%% Using    C IV

%% In [6]: log10((10**47.216)/((1.26e38)*(10**9.49)))                                                                                                                                             
%%Out[6]: -0.37437054511756207
%%
%%In [7]: log10((10**46.166)/((1.26e38)*(10**8.742)))                                                                                                                                            
%%Out[7]: -0.676370545117567
%%
%%In [8]: log10((10**46.721)/((1.26e38)*(10**9.1289)))                                                                                                                                           
%%Out[8]: -0.5082705451175662
%%
%%In [9]: log10((10**46.231)/((1.26e38)*(10**8.733)))                                                                                                                                            
%%Out[9]: -0.6023705451175618

\begin{table*}
  \begin{tabular}{l l  ll ll cc r}
    \hline
    \hline
\multirow{ 2}{*}{Object}  & \multirow{ 2}{*}{MJD} & \multirow{ 2}{*}{$M_i$} &  \multirow{ 2}{*}{$L_{\rm bol}$} & \multicolumn{2}{c}{$M_{\rm BH}$}     & \multicolumn{2}{c}{$\eta_{\rm Edd}$} &  Ref. \\
                                      &                                   &                                      &                                                 & \mgii                  & \civ                    &  \mgii           &  \civ                       &        \\
    \hline
                                     & 53498                         & -27.74                          & 47.216$\pm$0.004                 &  9.55$\pm$0.05  & 9.49$\pm$0.04   &     -0.434    & -0.374                   & Shen2011\\
 J1205+3422                & 58538                         &                                      &                                                 &                             &                             &                    &                               & \\
                                     & 58693                         &                                     &                                                  &                             &                             &                    &                               &   \\
    \hline 
                                    & 54553                          & -26.75                         & 46.166$\pm$0.04                    & 9.03$\pm$0.37    & 8.74$\pm$0.15  &  -0.964      & -0.677                      & Shen2011\\
 J1638+2827                & 55832                         &  -26.40                         & 46.721$\pm$0.073                 & 9.34                      &  9.13                    &  -0.717     & -0.509                          & Kol017\\
                                    & 58583                          &                                      &                                                &                               &                            &                   &                                      &  \\
    \hline 
                                   & 56189                          &  -25.46                         & 46.231$\pm$0.073                 &  9.45                     & 8.73                     &  -1.317   & -0.602                            & Kol017\\
J2228+2201               & 56960                          &                                       &                                                &                              &                               &                 &                                        & \\
                                  & 58693                          &                                       &                                               &                               &                                &                &                   &  \\
    \hline
    \hline
  \end{tabular}
  \caption{$\eta_{\rm Edd}$ is the base 10 logarithm of the Eddington ratio.
        Shen11 is \citet{Shen2011}. 
        Kol17 is \citet{Kozlowski2017}.
        Measured line values using the methods and catalogue from
\citet{Shen2011}.  $M_{\rm I}(z=2)$ is the Absolute $i$-band magnitude
$K$-corrected to $z = 2$; Bolometric luminosity computed from the
monochromatic luminosity at 1350\AA\ using the spectral fits and
bolometric corrections (BC = 3.81) in \citet{Richards2006b}; Line
luminosity, FWHM, rest-frame equivalent width, and their errors for
the whole \civ\ profile.  Power-law slope $\alpha_{\lambda}$ for the
continuum fit for \civ; Virial BH masses using calibrations of
\citet{VestergaardPeterson2006}.  Eddington ratio computed using the
virial BH mass.
      }
\end{table*}

@@

\subsection{The Baldwin Effect}
The variable properties of the rest-frame UV quasar emission lines
have been long studied, with the global (or ensemble) Baldwin Effect
(the anti-correlation between the EW of the emission line and the
underlying continuum luminosity of single-epoch observations of a
large number of AGN, first noted in \citet{Baldwin1977}.  More
recently, the intrinsic Baldwin effect, the same anti-correlation but
in an individual, variable AGN.  The variable properties of high-$z$
quasars, however, has been less well studied. Reasons for this include
more massive systems will tend to have longer timescales e.g. with
$R_{\rm Sch}= 2GM /c^{2}$ and $t_{\rm dyn} =\sqrt{ R^{3}/GM}$ means
$t_{\rm dyn} \sim 6 G^{3} M^{3} / GM c^{2} \sim 6G^{2} M^{2} / c^{2}
$.

\subsection{Eddington ratios and State Changes} 
The broad UV and optical lines in quasars are most sensitive to the extreme ultraviolet (EUV)
part of the spectral energy distribution (SED), with \civ\ (and indeed \heii and \nv) being at
the higher energy end of the EUV distribution. 
\citet{NodaDone2018} note that The soft X-ray excess produces most of
the ionizing photons, so its dramatic drop leads to the disappearance
of the broad-line region, driving the ``changing-look'' phenomena.
major difference is that radiation pressure should be much more
important in AGNs, so that the sound speed is much faster than
expected from the gas temperature.
%%
This spectral hardening appears similar to the soft-to-hard state
transition in black hole binaries at $L / L_{\rm Edd} \sim 0.02$
(i.e. $\eta_{\rm Edd} \sim -1.7$, where the inner disc evaporates into
an advection dominated accretion flow, while the overall drop in
luminosity appears consistent with the hydrogen ionization disc
instability.  Crucially \citet{NodaDone2018} make the prediction that
all changing-look AGNs are similarly associated with the state
transition at $L / L_{\rm Edd} \sim$a few per cent.

@@

Etiam mollis viverra nisi eget aliquet. Aliquam erat volutpat. Vivamus
tristique, nisl eu malesuada semper, libero tortor convallis elit, a
scelerisque orci nisi lacinia turpis. In lacinia ultrices
volutpat. Proin ultrices luctus tellus, in placerat eros tincidunt
id. Ut varius iaculis quam in consequat. Nulla nec orci est, sit amet
Aliquam ac metus nec odio tempus pharetra sed nec diam. Sed eget arcu
nulla. Etiam elementum ultrices ligula, at iaculis libero feugiat
bibendum. Suspendisse potenti. Nam pharetra adipiscing
euismod. Quisque imperdiet dignissim odio, sed volutpat justo
tincidunt eu. Nunc vehicula pharetra suscipit. Integer aliquet pretium
ipsum vel ultrices. Nam rutrum nibh ac quam pulvinar molestie.

### 2019-11-07 tex/other_tex/201910/CIV_CLQs_20191022.tex
CIV tex/ tidy up

\begin{abstract}
We report on three redshift $z>2$ quasars with dramatic changes in
their \civ\ emission lines, the first `Changing-Look'' quasars at high
redshift.  This is also the first time the changing-look behaviour has
been seen in a high-ionization emission line.
%%
SDSS J1205+3422, J1638+2827 and J2228+2201 show interesting behaviour
in their observed optical light curves, and subsequent spectroscopy
shows significant changes in the \civ\ broad emission line, with both
line collapse and emergence being displayed in rest-frame timescales
of $\sim$240-1640 days.
%%
Where observed, the profile of the Ly$\alpha$/\nv emission complex
also changes, and there is tentative evidence for changes in the \mgii
line.
%%
Although line measurements from the three quasars show large changes
in the \civ\ line flux-line width plane, the quasars are not seen to
be outliers when considered against the full $z>2$ quasar population
in terms of (rest) Equivalent Width and FWHM properties.
%%
We put these observations in context with recent ``state-change''
models, but note that even in their `low-state', the \civ\ CLQs are
above $\sim10\%$ in Eddington luminosity.
\end{abstract}

@@

  \begin{tabular}{r  r  r r r   r r r r}

@@

    J120544.7+342252.4   & 2.068$\pm$0.0003         &   18.27             &  53498   &  2005-May-08   & SDSS             & tbd                          & 2089-427             & \\

@@

    J163852.9+282707.7   & 2.185$\pm$0.0004       &   19.77              &  54553    & 2008-Mar-28     & SDSS             & tbd                          &  2948-614              & \\
    249.720558	                 &  2.186$\pm$0.0007     &                          &  55832    & 2011-Sep-28     & BOSS            &   3603                  &  5201-178            & \\

@@

    J222818.7+220102.9   & 2.217$\pm$0.0021      & 19.97                &  56189    & 2012-Sep-19      & BOSS             &  2702                 &   6118-720          & \\
    337.078194                 & 2.222$\pm$0.0004      &                          &  56960    & 2014-Oct-30      & BOSS             & tbd                         &   7582-790          & QSO1-REOBS \\ 

@@

    \rowfont{\color{blue}}       & 53498    & \civ      &  523$\pm$14    &   9816$\pm$321        &  1407$\pm$81      &  16.83$\pm$0.43    & $-1.64\pm0.01$      &           QSFit  \\

@@

   \rowfont{\color{blue}}          & 54553     & \civ      & 257$\pm$13    &  4550$\pm$371     &   55$\pm$450         &  31.80$\pm$1.51     &                                &  $^{(d)}$QSFit  \\
   \rowfont{\color{blue}}          & 54553     & \civ      &  281                  &    4181$\pm$736   &    680$\pm$133       &    100.5$\pm$11.9  & $-0.14\pm0.68$    &  Shen2011   \\

@@

                                              & 56960         & \lya    & 121$\pm$38     & 3580$\pm$673       &     -16$\pm$171     &  13.52 $\pm$4.21    &                                  &   QSFit   \\  

@@

           & \textcolor{blue}{56960} & \textcolor{blue}{\civ}     & \textcolor{blue}{254$\pm$5}    &  \textcolor{blue}{8731$\pm$172}   &   \textcolor{blue}{92$\pm$68}  &   \textcolor{blue}{45.35$\pm$0.90}    &    &    \\   

>>>>

\begin{abstract}
We report on three redshift $z>2$ quasars with dramatic changes in
their \civ\ emission lines, the first `Changing-Look'' quasars at high
redshift.  This is also the first time the changing-look behaviour has
been seen in a high-ionization emission line.
%%
SDSS J1205+3422, J1638+2827 and J2228+2201 show interesting behaviour
in their observed optical light curves, and subsequent spectroscopy
shows significant changes in the \civ\ broad emission line, with both
line collapse and emergence being displayed in rest-frame timescales
of $\sim$240-1640 days.
%%
Where observed, the profile of the Ly$\alpha$/\nv emission complex
also changes, and there is tentative evidence for changes in the \mgii
line.
%%
Although line measurements from the three quasars show large changes
in the \civ\ line flux-line width plane, the quasars are not seen to
be outliers when considered against the full $z\sim2$ quasar population
in terms of (rest) Equivalent Width and FWHM properties.
%%
We put these observations in context with recent ``state-change''
models, but note that even in their `low-state', the \civ\ CLQs are
above $\sim$10\% in Eddington luminosity.
\end{abstract}

@@

  \begin{tabular}{r  l  r r r   r r r r}

@@

    J120544.7+342252.4   & 2.068$\pm$0.0003         &   18.27             &  53498   &  2005-May-08   & SDSS             &                            & 2089-427             & \\

@@

    J163852.9+282707.7   & 2.185$\pm$0.0004       &   19.77              &  54553    & 2008-Mar-28     & SDSS           &                               &  2948-614              & \\
    249.720558	                 &  2.186$\pm$0.0007     &                          &  55832    & 2011-Sep-28     & BOSS            &   3600                  &  5201-178            & \\

@@

    J222818.7+220102.9   & 2.217$\pm$0.0021      & 19.97                &  56189    & 2012-Sep-19     & BOSS             &  2700                 &   6118-720          & \\
    337.078194                 & 2.222$\pm$0.0004      &                          &  56960    & 2014-Oct-30     & BOSS             &                             &   7582-790          & QSO1-REOBS \\ 

@@

    \rowfont{\color{blue}}       & 53498    & \civ      &  511$\pm$13    &   9593$\pm$287        &  1172$\pm$74      & 15.9$\pm$0.4      & $-1.64\pm0.01$      &           QSFit  \\

@@

   \rowfont{\color{blue}}          & 54553     & \civ      & 257$\pm$13    &  4550$\pm$371     &   55$\pm$450         &   31.8$\pm$1.5     &                                &  $^{(d)}$QSFit  \\
   \rowfont{\color{blue}}          & 54553     & \civ      &  281                  &    4181$\pm$736   &    680$\pm$133       & 100.5$\pm$12  & $-0.14\pm0.68$    &  Shen2011   \\

@@

                                              & 56960         & \lya    & 121$\pm$38     & 3580$\pm$673       &     -16$\pm$171     &  13.52 $\pm$4.21    & ---                           &   QSFit   \\  

@@

           & \textcolor{blue}{56960} & \textcolor{blue}{\civ}  & \textcolor{blue}{229$\pm$4}  &  \textcolor{blue}{8911$\pm$169}   &   \textcolor{blue}{92$\pm$68}  &   \textcolor{blue}{46.8$\pm$0.9}    & ---    & \textcolor{blue}{QSFit}     \\   
